Decoupled Fallback: A Portable Single-Pass GPU Scan

### 2025-02-06 paper.tex
Add Real Data

\section{Evaluation}
\subsection{Experimental Setup}
We configure workgroups with a size of 256 threads, where each thread processes 4 vectors of size 4, resulting in 16 elements per thread and a total work tile size of 4096. We find a maximum spin count of 4 as offering good responsiveness to blocking, while minimizing false-positives fallbacks. As the size of the local spine is inversely proportional to the subgroup size, we allocate sufficient shared memory to accommodate the minimum subgroup size specified by the WebGPU standard, which is 4. Despite the additional shared memory required for conflict avoidance and fallback operations, total shared memory usage remains modest at only 1 KB. For all tests, we use an input size of $2^{25}$.

Something about devices here.
\subsection{Performance Comparison}
To evaluate the performance of \emph{Decoupled-Fallback}, we compare it against two key benchmarks: \emph{Reduce-Then-Scan} (RTS) and a \emph{Memcopy} operation. RTS serves as a baseline to demonstrate the performance gains of our approach over the previous FPG-constrained technique. Memcopy is included as a point of reference, as it shares the same $2n$ global memory movement required by scan operations.

\thomas{Forgot about tables. Mega table incoming.}

\thomas{Performance is good}

@@

\subsection{Emulated Blocking}
To demonstrate the resilience of our technique under fallback rates significantly higher than those observed organically on devices, we intentionally force fallbacks by omitting the posting of work tile reductions. Starting with infrequent omissions---1 in 512 workgroups---we progressively increase the rate to near-continuous levels, reaching 1 in 2 workgroups.
\begin{figure}
  \centering
  \begin{subfigure}{.9\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/fallbacksInitiated_plot.pdf}
    \caption{Fallbacks Initiated}
    \label{fig:fallbacks_initiated}
  \end{subfigure}
  \begin{subfigure}{.9\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/totalSpins_plot.pdf}
    \caption{Total Spins}
    \label{fig:total_spins}
  \end{subfigure}
  \begin{subfigure}{.9\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/lookbackLength_plot.pdf}
    \caption{Lookback Length}
    \label{fig:lookback_length}
  \end{subfigure}
  \begin{subfigure}{.9\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/time_plot.pdf}
    \caption{Speed vs Memcopy}
    \label{fig:execution_time}
  \end{subfigure}
  \caption{Simulated Blocking\john{Text in these plots---axis labels, etc.---needs to be larger. If it's significantly smaller than body text, which it is, it's just not legible.}}
  \label{fig:vertical_images}
\end{figure}

>>>>

\section{Evaluation}
\subsection{Experimental Setup}
\begin{table*}
  \centering
  \begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} l r r r r r}
    \toprule
    Metric & Nvidia 2080 Super & Intel HD620 & AMD 7900 XT & ARM Mali-G78 MP20 & Apple M1 Pro Max \\  
    \midrule
    Memcopy                  & 54.203  & 1.498   & 71.190  & 106.663 & 54.203  \\  
    RTS                      & 35.709  & 1.060   & 62.868  & 82.302  & 35.709  \\  
    CSDLDF                   & 53.256  & 1.487   & 95.737  & 112.430 & 53.256  \\  
    CSDLDF/Memcopy           & 0.983   & 0.992   & 1.345   & 1.054   & 0.983   \\  
    CSDLDF/RTS               & 1.491   & 1.402   & 1.523   & 1.366   & 1.491   \\  
    Avg. Spins/Workgroup     & 0.418   & 0.306   & 0.133   & 1.380   & 0.418   \\  
    Avg. Lookback/Workgroup  & 13.139  & 1.725   & 20.867  & 3.510   & 13.139  \\  
    Fallbacks Initiated      & 88.384  & 260.364 & 10.127  & 2202.045& 88.384  \\  
    Successful Insertions    & 1.205   & 6.084   & 1.001   & 57.875  & 1.205   \\  
    \bottomrule
  \end{tabular*}
  \caption{Harmonic Means and Ratios for Various Devices, (Mali Absolute Statistics incorrect, need M1 Statistics)}
  \label{tab:results}
\end{table*}

We configure workgroups with a size $w = 256$ threads, where each thread processes $p = 4$ vectors of size 4, resulting in 16 elements per thread and a total work tile size $t = 4096$. We find a maximum spin count of 4 as offering good responsiveness to blocking, while minimizing false-positives fallbacks. As the size of the local spine is inversely proportional to the subgroup size, we allocate sufficient shared memory to accommodate the minimum subgroup size specified by the WebGPU standard, which is 4. Due to our overall thriftiness with shared memory, total shared memory usage remains modest at only 1 KB. We choose a diverse range of hardware vendors, with. For all tests, we use an input size of $2^{25}$.

\subsection{Performance Comparison}
To evaluate the performance of \emph{Decoupled-Fallback}, we compare it against two key benchmarks: \emph{Reduce-Then-Scan} (RTS) and a \emph{Memcopy} operation. RTS serves as a baseline to demonstrate the performance gains of our approach over the previous FPG-constrained technique. Memcopy is included as a point of reference, as it shares the same $2n$ global memory movement required by scan operations.

@@

\subsection{Emulated Blocking}
To demonstrate the resilience of our technique under fallback rates significantly higher than those observed organically on devices, we intentionally force fallbacks by omitting the posting of work tile reductions. Starting with infrequent omissions---1 in 512 workgroups---we progressively increase the rate to near-continuous levels, reaching 1 in 2 workgroups.
\begin{figure*}[t] % [t] forces the figure to appear at the top
  \centering
  % First row: two subfigures side by side
  \begin{subfigure}[t]{0.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/fallbacksInitiated_plot.pdf}
    \caption{Fallbacks Initiated}
    \label{fig:fallbacks_initiated}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/totalSpins_plot.pdf}
    \caption{Total Spins}
    \label{fig:total_spins}
  \end{subfigure}

  \vspace{0.2em} % Adjust vertical spacing between rows

  % Second row: two subfigures side by side
  \begin{subfigure}[t]{0.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/lookbackLength_plot.pdf}
    \caption{Lookback Length}
    \label{fig:lookback_length}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.48\linewidth}
    \centering
    \includegraphics[width=\linewidth]{graphics/time_plot.pdf}
    \caption{Throughput vs Baseline}
    \label{fig:execution_time}
  \end{subfigure}

  \caption{Simulated Blocking}
  \label{fig:vertical_images}
\end{figure*}
